---
title: "White Wine Quality by Xuan Guo"
author: "Xuan Guo"
date: "8/2/2017"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(GGally)
library(reshape2)
library(dplyr)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Load the Data
WhiteWine = read.csv('wineQualityWhites.csv')[,-1]

```

# Introduction

This project is an exploratory data analysis on the white wine quality dataset,
aiming to figure out how the input variables relate with the output variable, 
that is, to find out how white wine quality is effected by the chemicals. The 
white wine quality data is a tidy dataset with 4898 obeservations and 11 
input variables--chemicals, and one output variable -- quality. 

# Attribute Information:

Input variables (based on physicochemical tests):

1 - fixed acidity (tartaric acid - g / dm^3)

2 - volatile acidity (acetic acid - g / dm^3)

3 - citric acid (g / dm^3)

4 - residual sugar (g / dm^3)

5 - chlorides (sodium chloride - g / dm^3

6 - free sulfur dioxide (mg / dm^3)

7 - total sulfur dioxide (mg / dm^3)

8 - density (g / cm^3)

9 - pH

10 - sulphates (potassium sulphate - g / dm3)

11 - alcohol (% by volume)

Output variable (based on sensory data):

12 - quality (score between 0 and 10)

New input variables:

13 - acid(g / dm^3)

14 - bound sulfur dioxide(mg / dm^3)

15 - total(g / dm^3)

16 - free.sulfur.ratio(range from 0 to 1)

New output variables:

17 - quality.int(int version of quality)

18 - quality.bucket((0,3],(3,5],(5,7])

Description of attributes:

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not
evaporate readily)

2 - volatile acidity: the amount of acetic acid in wine, which at too high of 
levels can lead to an unpleasant, vinegar taste

3 - citric acid: found in small quantities, citric acid can add 'freshness' and 
flavor to wines

4 - residual sugar: the amount of sugar remaining after fermentation stops, it's
rare to find wines with less than 1 gram/liter and wines with greater than 45 
grams/liter are considered sweet

5 - chlorides: the amount of salt in the wine

6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between 
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial 
growth and the oxidation of wine

7 - total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO2 is mostly undetectable in wine, but at free SO2 
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

8 - density: the density of water is close to that of water depending on the 
percent alcohol and sugar content

9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) 
to 14 (very basic); most wines are between 3-4 on the pH scale

10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02)
levels, wich acts as an antimicrobial and antioxidant

11 - alcohol: the percent alcohol content of the wine

Output variable (based on sensory data): 

12 - quality (score between 0 and 10)
New feature descripstions are listed in the questions for univariable plots 
section.


# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Add_New_Features}

#Add new features:
WhiteWine$acid = WhiteWine$fixed.acidity + WhiteWine$volatile.acidity +
                  WhiteWine$citric.acid
WhiteWine$bound.sulfur.dioxide = WhiteWine$total.sulfur.dioxide - 
                  WhiteWine$free.sulfur.dioxide
WhiteWine$total = WhiteWine$acid + WhiteWine$residual.sugar + WhiteWine$chlorides +
                  WhiteWine$total.sulfur.dioxide / 1000 +
                  WhiteWine$sulphates
WhiteWine$free.sulfur.ratio = WhiteWine$free.sulfur.dioxide /
                               WhiteWine$total.sulfur.dioxide

WhiteWine$quality = factor(WhiteWine$quality)
WhiteWine$quality.int = as.numeric(WhiteWine$quality)
WhiteWine$quality.bucket = cut(WhiteWine$quality.int, breaks=c(0,3,5,7))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Summary_of_data}

summary(WhiteWine)

```

This dataset has numeric variables like fixed acidity, volatile acidity and so 
on, with quality the only dummy variable. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
#create_plot function is used to reduce repetition
create_plot <- function(variable, binwidth, log = TRUE) {
  if (log == TRUE)
  {
    return(ggplot(aes_string(x = variable), data = WhiteWine) + 
           geom_histogram(binwidth = binwidth) +
           scale_x_log10() +
           xlab(paste('log10(', variable,')')))
  }
  else
  {
    return(ggplot(aes_string(x = variable), data = WhiteWine) + 
           geom_histogram(binwidth = binwidth))
  }
}
create_plot('fixed.acidity', binwidth = 0.20, FALSE)
  
```

The histogram of fixed.acidity shows that the feature approximately follows
normal distribution.  

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 = create_plot('volatile.acidity',0.03,FALSE)
p2 = create_plot('volatile.acidity',0.06,TRUE)
grid.arrange(p1, p2, ncol = 2)
WhiteWine$volatile.acidity = log10(WhiteWine$volatile.acidity)

```

The histogram of volatile.acidity shows that the feature approximately follows
normal distribution, but it is slightly right-skewed. Therefore, loo tranform 
the volatile data.

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('citric.acid', 0.02, FALSE)

```

The histogram of citric.acidity shows that the feature approximately follows
normal distribution, except that it has several unexpected spikes.  

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 = create_plot('residual.sugar',3,FALSE)
p2 = create_plot('residual.sugar',0.1,TRUE)
grid.arrange(p1,p2,ncol = 2)
WhiteWine$residual.sugar = log10(WhiteWine$residual.sugar)

```

Since residual.sugar is highly right-skewed, I log-transformed it. The adjusted
histogram shows that log(residual.sugar) approximately follows a bimodal normal 
distribution. 
Therefore, change the original residual.sugar into log(residual.sugar) in the 
dataset. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('chlorides', 0.005, FALSE)

```

The histogram of chlorides shows that the feature approximately follows
normal distribution. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('free.sulfur.dioxide', 3, FALSE)

```

The histogram of free.sulfur.dioxide shows that the feature approximately 
follows normal distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('total.sulfur.dioxide', 10, FALSE)

```

The histogram of total.sulfur.dioxide shows that the feature approximately 
follows normal distribution.


```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('density', 0.0008, FALSE)

```

The histogram of density shows that the feature approximately follows normal
distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('pH', 0.05, FALSE)

```

The histogram of pH shows that the feature nicely follows
normal distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('sulphates', 0.03, FALSE)

```

The histogram of sulphates shows that the feature approximately follows
normal distribution although slightly right-skewed. 


```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('alcohol', 0.2, FALSE)

```

The histogram of alcohol shows that the feature has multiple spikes.

```{r echo=FALSE, message=FALSE, warning=FALSE}

qplot(x = quality, data = WhiteWine)

```

As we can see, there are more quality 6 whitewine than any other other quality 
wine in the dataset.


```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('acid', 0.15, FALSE)

```

The histogram of acid shows that the feature approximately follows normal
distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('bound.sulfur.dioxide', 5, FALSE)

```

The histogram of bound.sulfur.dioxide shows that the feature approximately 
follows normal distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 = create_plot('total', 1.5, FALSE)
p2 = create_plot('total', 0.02, TRUE)
grid.arrange(p1, p2, ncol = 2)
WhiteWine$total = log10(WhiteWine$total)

```

Since the feature total is highly right-skewed, I log transformed it the 
dataset. The distribution of the transformed data seems to contain multiple
gaussian distributions.

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_plot('free.sulfur.ratio', 0.015, FALSE)

```

The histogram of free.sulfur.ratio shows that the feature approximately 
follows normal distribution.

# Univariate Analysis

### What is the structure of your dataset?
The dataset has 4898 samples, with 11 variables numeric and quality, the only
discrete variable. After creating 6 new features, I made the dataset 18 
variables.

### What is/are the main feature(s) of interest in your dataset?

The relationships between the possible predictors(variables except for quality)
and the output(quality).

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The relationships between the input features. For example, the relationship 
between density and total might be helpful in figuring out how chemicals in the
feature list would influence the quality. 

### Did you create any new variables from existing variables in the dataset?
Yes. 
I created six new variables.

(1)acid: the sum of fixed.acidity, volatile.acidity and citric acidity. Since 
there are three types of acid listed in the dataset, it maybe helpful to see the 
influence of the total amount of acid.

(2)bound.sulfur.dioxide: it equals to total.sulfur.dioxide minus 
free.sulfur.dioxide. Since total.sulfur.dioxide is the sum of 
free.sulfur.dioxide and bound.sulfur.dioxide, it would be interesting to see the
relationship between these three.

(3)total: the sum of acid,residual.sugar, chlorides, total.sulfur.dioxide, 
sulphates. I created total because I wanted to figure out the connection between
all the substances in the predictor list and density.

(4)free.sulfur.ratio: the ratio of free.sulfur.dioxide and total.sulfur.dioxide.
I created this feature because from plot three in the final section the ratio of
free.sulfur.dioxide and total.sulfur.dioxide changes according to quality.

(5)quality.int: the digit version of quality. I added this variable so that a 
correlation analysis would be possible bwtween quality and other variables.

(6)quality.bucket: Since there exists too many values for quality variable which 
makes it hard to explore how other variables relates with quality, it would be 
helpful to divide the quality into buckets. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Yes. The residual sugar data is highly right-skewed and I log-transfered it so 
that it is approxiamtely a bimodal distribution, with the first peak after 1 and
the second around 10. I aslo log scaled 'total' because it's highly skewed.

# Bivariate Plots Section

The correlation matrix of the whole dataset is shown below.

```{r, echo=FALSE,message=FALSE, warning=FALSE}

temp_WhiteWine = subset(WhiteWine, select = -c(quality,quality.bucket))
cor_matrix = cor(temp_WhiteWine)
cor_matrix

```


```{r echo=FALSE,message=FALSE, warning=FALSE, List_all_the_Greater_than_0.6_correlations}

for (i in (1: nrow(cor_matrix)))
{
  for (j in (1: ncol(cor_matrix)))
  {
    if (j >= i)
    {
      break
    }
    value = cor_matrix[i,j]
    if (value > 0.6 | value < -0.6)
    {
      value = signif(value,digits=5)
      temp = paste(row.names(cor_matrix)[i],'and',colnames(cor_matrix)[j],'have correlation of', value)
      print(temp)
    }
  }
}

```

Explore the highly correlated variables below.

## Between Supporting Variabels

```{r echo=FALSE, message=FALSE, warning=FALSE}
bi_plt1 = ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), data = WhiteWine) + 
          geom_jitter(alpha=0.5)+
          coord_cartesian(xlim=c(quantile(WhiteWine$total.sulfur.dioxide,0.01),
                                quantile(WhiteWine$total.sulfur.dioxide,0.99)))+
          geom_smooth(method = 'gam')
bi_plt2 = ggplot(aes(x = density, y = residual.sugar), data = WhiteWine) + 
          geom_jitter(alpha = 0.5) +
          coord_cartesian(xlim=c(quantile(WhiteWine$density,0.01),
                                  quantile(WhiteWine$density,0.99))) +
          geom_smooth(method = 'gam')
bi_plt3 = ggplot(aes(x = acid, y = fixed.acidity), data = WhiteWine) + 
          geom_jitter(alpha = 0.5) +
          coord_cartesian(xlim=c(quantile(WhiteWine$acid,0.01),
                                  quantile(WhiteWine$acid,0.99))) +
          geom_smooth(method = 'gam')
bi_plt4 = ggplot(aes(x = bound.sulfur.dioxide, y = total.sulfur.dioxide), data = WhiteWine) + 
          geom_jitter(alpha = 0.5) +
          coord_cartesian(xlim=c(quantile(WhiteWine$bound.sulfur.dioxide,0.01),
                                  quantile(WhiteWine$bound.sulfur.dioxide,0.99))) +
          geom_smooth(method = 'gam')
bi_plt5 = ggplot(aes(x = total, y = residual.sugar), data = WhiteWine) + 
          geom_jitter(alpha = 0.5) +
          coord_cartesian(xlim=c(quantile(WhiteWine$total,0.01),
                                  quantile(WhiteWine$total,0.99))) +
          geom_smooth(method = 'gam')
bi_plt6 = ggplot(aes(x = total, y = density), data = WhiteWine) + 
          geom_jitter(alpha = 0.5) +
          coord_cartesian(xlim=c(quantile(WhiteWine$total,0.01),
                                  quantile(WhiteWine$total,0.99))) +
          geom_smooth(method = 'gam')
bi_plt7 = ggplot(aes(x = alcohol, y = density), data = WhiteWine) + 
          geom_jitter(alpha = 0.5) +
          coord_cartesian(xlim=c(quantile(WhiteWine$alcohol,0.01),
                                  quantile(WhiteWine$alcohol,0.99))) +
          geom_smooth(method = 'gam')

grid.arrange(bi_plt1, bi_plt2, bi_plt3, bi_plt4, bi_plt5, bi_plt6, bi_plt7,
             ncol = 2)

```

As we can see from the figure, total and residual.sugar, total.sulfur.dioxide
and free.sulfur.dioxide, acid and fixed.acidity, density and residual.sugar are 
highly correlated. The results also indicates that

(1)There are generally more fixed acidity than citric acid and volatile acid;

(2)White wine in the sample usually contains more residual sugar than other
chemicals in the dataset. Since feature residual.sugar contributes a lot to 
density and total;

(3)Generally, more bound sulfur dioxide exists in sample white wine than free 
sulfur dioxide because the bound one is more correlated with total sulfur
dioxide.

#Between Supporting Variabels And Quality

```{r echo=FALSE, message=FALSE, warning=FALSE}

#Correlations with quality
cor_matrix[,'quality.int']

```

As we see from correlation matrix, alcohol, density, bound.sulfur.dioxide, 
chlorides, volatile.acidity, free.sulfur.ratio are more correlated with quality 
than other features.

Explore those variables below:


```{r echo=FALSE, message=FALSE, warning=FALSE}

bi_plt8 = ggplot(aes(x = quality, y = alcohol), data = WhiteWine) +
  geom_boxplot()
bi_plt9 = ggplot(aes(x = quality, y = density), data = WhiteWine) +
  geom_boxplot()
bi_plt10 = ggplot(aes(x = quality, y = bound.sulfur.dioxide), data = WhiteWine) +
  geom_boxplot()
bi_plt11 = ggplot(aes(x = quality, y = chlorides), data = WhiteWine) +
  geom_boxplot()
bi_plt12 = ggplot(aes(x = quality, y = volatile.acidity), data = WhiteWine) +
  geom_boxplot()
bi_plt13 = ggplot(aes(x = quality, y = free.sulfur.ratio), data = WhiteWine) +
  geom_boxplot()
grid.arrange(bi_plt8,bi_plt9,bi_plt10,bi_plt11,bi_plt12,bi_plt13,ncol = 2)

```

From the plots above, we could see that 

(1)Alcohol is much more varied on quality than other features. It has the 
greatest correlation(0.435574715) with quality among all the supporting 
variables.

(2)Chlorides and volatile both have many outliers: Chlorides has the most 
outliers on quality 5 and 6; Volatile acidity has the most outliers on quality
5 and 6.

```{r echo=FALSE, message=FALSE, warning=FALSE}

cor.test(WhiteWine$free.sulfur.ratio, WhiteWine$quality.int)

quality_groups = group_by(WhiteWine, quality.bucket)
WhiteWine.sulfur_by_quality = WhiteWine%>%
                              group_by(quality.bucket)%>%
                              summarise(sul_median = median(free.sulfur.ratio))%>%
                              arrange(quality.bucket)

ggplot(aes(x = free.sulfur.ratio, fill =  quality.bucket), data = WhiteWine)+
  geom_density(alpha = 0.6) +
  geom_vline(aes(xintercept = sul_median, colour = quality.bucket), 
             data = WhiteWine.sulfur_by_quality,
             size = 1,
             linetype = 'dotdash')

```


The plot two will explain why the relationship between free.sulfur.ratio and 
quality is specially analyzed here. 
free.sulfur.ratio and WhiteWine$quality.int have a correlation of 0.1972. The 
The dotted lines are the medians of free sulfur ratio on each quality bucket. As
the plot shows, median of the free sulfur ratio is larger as quality is higher.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
Many supporting variables are highly related with each other, 6 pairs have 
correlation greater than 6(except for free.sulfur.ratio). Among all the 
correlations with quality, the greatest one is 0.4356. 

Alcohol and chlorides both have relatively great correlation with quantity. 
However, we could see from the figure that the median and quantile of alcohol 
both vary a lot on quantity; feature chlorides' median and quantile do not vary 
much on quantity.
The features of interest have greater correlation with quality than other 
features. They are alcohol, density, bound.sulfur.dioxide, chlorides and 
volatile.acidity.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
Yes.

The high correlations between total.sulfur.dioxide and free.sulfur.dioxide,
density and residual.sugar, acid and fixed.acidity, bound.sulfur.dioxide and total.sulfur.dioxide, total and residual.sugar, total and density, alcohol and
density indicate the quantity relationship between these chemicals. The greater
they are correlated, the similar pattern their quanities are.

### What was the strongest relationship you found?
'acid' and 'fixed.acidity'(0.98718)

# Multivariate Plots Section

```{r,echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide, colour = quality.bucket), 
                    data = WhiteWine) + 
          geom_jitter(alpha = 0.3) +
          coord_cartesian(ylim=c(0,
                                  quantile(WhiteWine$free.sulfur.dioxide,0.99))) +
          geom_smooth(method = 'loess')

```

The three lines have very different tangents. That's why I had the new feature
free.sulfur.ratio created and it has been analyzed in the bivariate section.


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = density, y = residual.sugar, colour = quality.bucket), 
          data = WhiteWine) + 
    geom_jitter(alpha = 0.3) +
    coord_cartesian(xlim = c(quantile(WhiteWine$density,0.01),
                             quantile(WhiteWine$density,0.99))) +
   geom_smooth(method = 'loess')


```

From the plot, we could see that the correlation of residual.sugar and density 
found in bivariate section still exists even on different quality values, but 
the slopes are slightly different on different quality.buckets.


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = acid, y = fixed.acidity, colour = quality.bucket), 
               data = WhiteWine) + 
        geom_jitter(alpha = 0.3) +
        coord_cartesian(xlim=c(quantile(WhiteWine$acid,0.01),
                                quantile(WhiteWine$acid,0.99))) +
        geom_smooth(method = 'loess')

```

From the plot, we could see that the correlation between acid and fixed.acidity
stays very the same on different values of quality.


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = bound.sulfur.dioxide, y = total.sulfur.dioxide,
                        colour = quality.bucket), data = WhiteWine) + 
          geom_jitter(alpha = 0.3) +
          coord_cartesian(xlim=c(quantile(WhiteWine$bound.sulfur.dioxide,0.01),
                                  quantile(WhiteWine$bound.sulfur.dioxide,0.99))) +
          geom_smooth(method = 'loess')

```

From the plot, we could see that the correlation between bound.sulfur.dioxide 
and total.sulfur.dioxide stays very the same on different values of quality.


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = total, y = residual.sugar, colour = quality.bucket), 
                    data = WhiteWine) + 
          geom_jitter(alpha = 0.3) +
          coord_cartesian(xlim=c(quantile(WhiteWine$total,0.01),
                                  quantile(WhiteWine$total,0.99))) +
          geom_smooth(method = 'loess')


```

From the plot, we could see that the correlation between total and  
residual.sugar differs a little on different value of quality.bucket. The slope
of the blue line is apparently increasing, while the slopes of the other two 
lines do not show an obvious rise. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = total, y = density, colour = quality.bucket),
                    data = WhiteWine) + 
          geom_jitter(alpha = 0.3) +
          coord_cartesian(xlim=c(quantile(WhiteWine$total,0.01),
                                        quantile(WhiteWine$total,0.99)),
                          ylim=c(quantile(WhiteWine$density,0.01),
                                        quantile(WhiteWine$density,0.99))) +
          geom_smooth(method = 'loess')

```

The slope of each trending line stays very the same when total is less than 8.
After total is greater than 8, the slope of (5,7](blue line) has an obvious rise
while the other two do not. The ratio of density and total changes a lot after
total reaches 8. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = alcohol, y = density, colour = quality.bucket),
                    data = WhiteWine) + 
          geom_jitter(alpha = 0.3) +
          coord_cartesian(xlim=c(quantile(WhiteWine$alcohol,0.01),
                                        quantile(WhiteWine$alcohol,0.99)),
                          ylim=c(quantile(WhiteWine$density,0.01),
                                        quantile(WhiteWine$density,0.99))) +
          geom_smooth(method = 'loess')
```

The correlation between alcohol and density stays very the same on each
quality.bucket.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

acids = WhiteWine[c('alcohol', 'acid', 'fixed.acidity', 'volatile.acidity', 
                    'citric.acid')]
acids.melt = melt(acids, id.vars = 'alcohol', variable.name = 'Acids')
ggplot(aes(x=alcohol, y = value, color = Acids), 
       data = acids.melt) +
       geom_jitter(alpha = 0.1) + 
       geom_smooth(method = 'loess')

```

Since alcohol is most correlated with quality, set it as x-axis. This plot
shows the comparison among different types of acids. We can see that acid and 
fixed.acidity have similar trends. volatile.acidity and citric.acidity have
similar trends.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

densities = WhiteWine[c('alcohol','residual.sugar','total')]
densities.melt = melt(densities, id.vars = 'alcohol', variable.name = 'features')
ggplot(aes(x=alcohol, y = value, color = features), 
       data = densities.melt) +
       geom_jitter(alpha = 0.1) + 
       geom_smooth(method = 'loess')

```

As showed in bivariate plot section, residual.sugar and total are highly 
correlated. They have similar trends but the trends seem to be very different when
alcohol is less than 12.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sulfur.dioxide = WhiteWine[c('alcohol', 'free.sulfur.dioxide',
                             'bound.sulfur.dioxide', 'total.sulfur.dioxide')]
sulfur.dioxide.melt = melt(sulfur.dioxide,id.vars = 'alcohol',
                           variable.name='sulfur.dioxide')
ggplot(aes(x=alcohol, y = value, color = sulfur.dioxide), 
       data = sulfur.dioxide.melt) +
       geom_jitter(alpha = 0.1) + 
       geom_smooth(method = 'loess')

```

This plot explores the quantity relationship among free sulfur dioxide, bound
sulfur dioxide, and total sulfur dioxide. Toal.sulfur.dioxide and 
bound.sulfur.dioxide have similar trends.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Yes. The positive correlatated features found in bivariate plot section still
strengthen each other when grouped by different value of quality. 

### Were there any interesting or surprising interactions between features?

The feature alcohol works well as an id when doing scatterplots, showing the 
quantities of different chemicals.

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

ggplot(aes(x = quality, y = alcohol), data = WhiteWine) +
  geom_boxplot() +
  ylab('alcohol (% by volume)') +
  ggtitle('Boxplot of Alcohol on Different Quality') +
  theme(plot.title = element_text(hjust = 0.5))

```

### Description One

This plot shows how alcohol differs with quality. Alcohol is most 
correlated with the output feature quality among all the supporting variables. 
The median of alcohol are increasing as quality increases from 5 to 9 and
decreasing as quality increases from 3 to 5.


### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide, 
                        colour = quality.bucket), data = WhiteWine) + 
          geom_jitter(alpha=0.3)+
          coord_cartesian(ylim=c(0,
                                quantile(WhiteWine$free.sulfur.dioxide,0.99))) +
          geom_smooth(method = 'loess') + 
          xlab('total sulfur dioxide (mg / dm^3)') + 
          ylab('free sulfur dioxide (mg / dm^3)') + 
          ggtitle('Scatterplot of Total Sulfur Dioxide and Free Sulfur Dioxide on Different Quality') + 
          theme(plot.title = element_text(hjust = 0.5))
  

```

### Description Two

This plot was used the verify the positive correlation on free sulfur dioxide 
and total sulfur dioxide when quality differs. It shows that the slope of each
line varies on quality bucket. The slope of the red line seems to increase after
total sulfur dioxide reaches 200, while the slope of the other two lines seems 
to be decreasing. Therefore, it would be helpful to analyze the ratio of free 
sulfur dioxide to the total one. 

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE,Plot_Three}

ggplot(aes(x = free.sulfur.ratio, fill =  quality.bucket), data = WhiteWine)+
  geom_density(alpha = 0.6) +
  geom_vline(aes(xintercept = sul_median, colour = quality.bucket), 
             data = WhiteWine.sulfur_by_quality,
             size = 1,
             linetype = 'dotdash') +
  xlab('Ratio of Free Sulfur Dioxide And Total Sulfur Dioxide') + 
  ylab('density') + 
  ggtitle('Density of Free Sulfur Dioxide Ratio on Different Quality') + 
  theme(plot.title = element_text(hjust = 0.5))

```

### Description Three

To analyze the ratio of free sulfur dioxide to total sulfur dioxide, I add a new
feature named 'free.sulfur.dioxide.ratio', which equals to free sulfur dioxide 
divided by total sulfur dioxide. The dashdotted lines are the medians of 'free.sulfur.dioxide.ratio' on different quality.bucket. As we can see from the 
plot, the median of the new feature is higher as the quality.bucket increases.

------

# Reflection

I found it hard to really explore any valuable things from the dataset at first.
There's no specific model that fits the data. The features are related but it's 
hard to find the pattern or trend without careful trials and observation. 
What surprised me was the relationship between the quality and the ratio of free 
sulfur dioxide and total sulfur dioxide. They are more positively related than 
I thought.
For future work, I think the EDA on WhiteWine dataset could be improved if the 
analysis could be relate more with some academic researches on white wine or 
chemical knowledge so that a comprehensive and systematical analysis could be 
conducted.


# Reference
https://archive.ics.uci.edu/ml/datasets/wine+quality
https://stackoverflow.com/questions/13035834/plot-every-column-in-a-data-frame-a
s-a-histogram-on-one-page-using-ggplot


